---
title: "gi-ws-09-1"
author: "Laura Giese, Johannes Schnell, Alena Schmid"
date: "22 Januar 2018"
output: html_document
---
###script for single tree detection with package LidR (for las data):

#---------------------------------------------------------------------------#
> structure: the scripts contain the following steps:

1. Set Environment
2. correct extent for U4775632
3. clip las data to requested extent (input here: two files)
4. generate normalized canopy height model (chm) (input: dtm)
5. clip chm to same extent as las data 
6. single tree detection with 3 diffenrent algorithms

#---------------------------------------------------------------------------#

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Set Environment

```{r, echo = T}
#---------------------------------------------------------------------------#
############ set environment and load packages  ########### 
#---------------------------------------------------------------------------#

setwd("C:/Users/Laura/Documents/Uni/Rmsc/Data/GIS/POINTCLOUD_processing/Einzelbaum/output")
# input directory with file with wrong extent
input = "C:/Users/Laura/Documents/Uni/Rmsc/Data/GIS/POINTCLOUD_processing/Einzelbaum/input/"
# output dir
output = "C:/Users/Laura/Documents/Uni/Rmsc/Data/GIS/POINTCLOUD_processing/Einzelbaum/output/"
#temp: input with file with corrected extent and second las file
temp = "C:/Users/Laura/Documents/Uni/Rmsc/Data/GIS/POINTCLOUD_processing/Einzelbaum/temp/"

# install packages
#install.packages("lidR", dependencies = T)
library(lidR)
library(raster)
library(rgdal)
library(maptools)
library(png)

## try http:// if https:// URLs are not supported
#source("https://bioconductor.org/biocLite.R")
#biocLite("EBImage")
#install.packages("EBImage")
#library(EBImage)
#unfortunatelly unable to load EBIMage package
```

#---------------------------------------------------------------------------#
####correct extent for U4775632

```{r, echo = T, eval = F}
#--------------------------------------------#
#######solve problem with missing extent######
#--------------------------------------------#
lasfile1 = paste0(input,"U4775632.las")
file.info(lasfile1)$size/1e6

las = readLAS(lasfile1, select = "xyz")
#use cut function to get right extent
subset = las %>% lasclipRectangle(xleft = 477000, ybottom = 5632000, xright = 478000, ytop = 5633000, inside = TRUE)

#kann raus
#ex_las=extent(477000, 478000, 5632000, 5633000)
#extent(las) = ex_las
#las@header@PHB$`Min X`<-as.numeric(c("477000"))
#las@header@PHB$`Min Y`<-as.numeric(c("5632000"))

#save corrected lasfile
writeLAS(subset, file = paste0(temp, "extU4775632.las"))
```

#---------------------------------------------------------------------------#
####START script for steps 3. to 6.

> structure of following part:

3. clip las data to requested extent (input here: two files)
4. generate normalized canopy height model (chm) (input: dtm)
5. clip chm to same extent as las data 
6. single tree detection with 3 diffenrent algorithms

```{r, echo = T}
#load las files: extU4775632.las with corrected extent and U4775631.las 
#from temp directory

#for loading las files
lasfile1 = paste0(input, "extU4775632.las")
lasfile2 = paste0(input, "U4775631.las")

#get info
file.info(lasfile1)$size/1e6
file.info(lasfile2)$size/1e6

#load lasfiles in one cloud
#make sure that input folder contains only the las files you need for your cloud
catalog = catalog(paste0(input))
las = readLAS(catalog, select = "xyz")

#change projection
las@crs@projargs<- c("+proj=utm +zone=32 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0+units=m +no_defs")
#las
#plot(las, color = "Z", colorPalette = heat.colors(50), bg = "black")


#--------------------------------------------#
#################clip las files###############
#--------------------------------------------#

#clip las files to extent 477425.0 477710.0 5631990.0 5632150.0 (nach Angabe von C. Reudenbach)
mrect = matrix(c(477425, 477710, 5631990, 5632150), ncol = 2) #create rect. matr
las_cl = las %>% lasclip("rectangle", mrect)
plot(las_cl, color = "Z", colorPalette = heat.colors(50), bg = "black", trim)
#writeLAS(las_cl, file = paste0(output, "las_clip.las"))

#--------------------------------------------#
#########normalize las file###################
#--------------------------------------------#

#create matrix for crop with shape file
coords = matrix(c(477425, 5632150,
                  477710, 5632150,
                  477710, 5631990,
                  477425, 5631990), ncol = 2, byrow = TRUE)

#create polygon shape file for crop dtm
P1 = Polygon(coords)
Ps1 = SpatialPolygons(list(Polygons(list(P1), ID = "a")), proj4string=CRS("+proj=utm +zone=32 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0+units=m +no_defs"))
plot(Ps1, axes = T)

#->load dtm (created in "gi-ws08-1 Lichtungen")
dtm = raster(paste0(input, "lichtung_GridSurf.asc"))
dtm@crs@projargs<- c("+proj=utm +zone=32 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0+units=m +no_defs")

#crop to same extent as las_cl
dtm_cr = crop(dtm, Ps1)
plot(dtm_cr)

#save cropped dtm
#writeRaster(dtm_cr, "dtm_crop_Baum.asc")

#normalize
lasnormalize(las_cl, dtm_cr)

#change coordinate system
las_cl@crs@projargs<- c("+proj=utm +zone=32 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0+units=m +no_defs")
las_cl
plot(las_cl)
#save clipped normalized las file
#writeLAS(las_cl, file = "las_clip_norm.las")

#----------------------------------------------------------------------#
###------------compute chm & single tree detection-------------------###
#----------------------------------------------------------------------#

# compute a canopy image
chm = grid_canopy(las_cl, res = 0.5, subcircle = 0.1, na.fill = "knnidw", k = 4)
chm = as.raster(chm)
# smoothing post-process (e.g. 2x mean)
kernel = matrix(1,3,3)
chm = raster::focal(chm, w = kernel, fun = mean)
chm = raster::focal(chm, w = kernel, fun = mean)
raster::plot(chm, col = height.colors(50)) # check the image

#----------------------------------------------------------------------#
# 1. way of segmentation: dalponte 
lastrees(las_cl, "dalponte2016", chm, th = 5)
#plot points that actually are trees
trees = lasfilter(las_cl, !is.na(treeID))
plot(trees, color = "treeID", colorPalette = random.colors(100))
#plot result
img <- readPNG("C:/Users/Laura/Documents/GitHub/msc-phygeo-class-of-2017-lgiese/gis/gi-ws-09-1/dalponte_st.png")
plot(1:1, type='n', xaxt="n", yaxt="n", xlab="", ylab = "", main = "dalponte")

lim <- par()
rasterImage(img, lim$usr[1], lim$usr[3], lim$usr[2], lim$usr[4])

#----------------------------------------------------------------------#
#2. way of segmentation: watershed
#doesn´t work in Rmd-file as EBImage package needs to be loaded from a special page on the internet?!
#see library(EBImage)
#lastrees(las_cl, "watershed", chm, th = 5)
# plot points that actually are trees
#trees = lasfilter(las_cl, !is.na(treeID))
#plot(trees, color = "treeID", colorPalette = random.colors(100))

#plot(trees, color = "treeID", colorPalette = random.colors(100))

#plot png instead
img <- readPNG("C:/Users/Laura/Documents/GitHub/msc-phygeo-class-of-2017-lgiese/gis/gi-ws-09-1/watershed_st.png")
plot(1:1, type='n', xaxt="n", yaxt="n", xlab="", ylab = "", main = "watershed")

lim <- par()
rasterImage(img, lim$usr[1], lim$usr[3], lim$usr[2], lim$usr[4])

#----------------------------------------------------------------------#
#3. way of segmentation: Li 2012
lastrees(las_cl, "li2012")
# plot points that actually are trees
trees = lasfilter(las_cl, !is.na(treeID))
plot(trees, color = "treeID", colorPalette = random.colors(100))
#plot
img <- readPNG("C:/Users/Laura/Documents/GitHub/msc-phygeo-class-of-2017-lgiese/gis/gi-ws-09-1/li2012seg_st.png")
plot(1:1, type='n', xaxt="n", yaxt="n", xlab="", ylab = "", main = "li2012")

lim <- par()
rasterImage(img, lim$usr[1], lim$usr[3], lim$usr[2], lim$usr[4])

#save las data with single trees
trees@crs@projargs<- c("+proj=utm +zone=32 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0+units=m +no_defs")
#writeLAS(trees, file = "sing_trees_dp.las")

```