---
title: "ws-05-2"
author: "Laura Giese, Johannes Schnell, Alena Schmid"
date: "5 Dezember 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Raw Script to calculate CH (normalized Canopy Height model) and HOME (Height Of Median Energy) with Fusion 
details: see Fusion Manual for details

> description: the script contains the following steps:

* set environment and read files
    + list input files
* preparations for CH
    + extract groundpoints of las data and create dtm
    + normalize las cloud 
* create height classes (slices of point cloud from zmin to zmax)
    + derive point density per height class (counting points per height class)
    + get .asc files with point density of every height class
* calculate Canopy height model
* calculate HOME and save it as .dtm file

```{r, eval = F, echo = T}
#---------------------------------------------------------------------------#
############ set environment and read files ########### 
#---------------------------------------------------------------------------#

setwd("C:/Users/Laura/Documents/Uni/Rmsc/Data/GIS/POINTCLOUD_processing")

mainDir <- "C:/Users/Laura/Documents/Uni/Rmsc/Data/GIS/POINTCLOUD_processing"
##input folder
inDir <- "/input/"
## output folder
outDir<- "/temp/"
## Fusion binary folder
Fusion<-"C:/FUSION/"

#list las files in input folder
las_files<-list.files(paste0(mainDir,inDir), pattern=".las$", full.names=TRUE,recursive = TRUE)
las_files
##write the whole file path and name into a .txt
lapply(las_files, write, paste0(mainDir,inDir,"lidar_files.txt"), append=T)

### basic informations
system(paste0(Fusion, "catalog.exe ",mainDir, inDir, "lidar_files.txt " ,mainDir, outDir, "info_caldern.html"))

#---------------------------------------------------------------------------#
########### preparations for CH ###########
#---------------------------------------------------------------------------#

### Create a .las with groundpoints only
system(paste0(Fusion, "clipdata.exe"," /class:2 ", mainDir, inDir, "lidar_files.txt ", 
              mainDir, outDir, "classified_GroundPts.las ", "476000 5630000 479000 5633000"))

### Create the required PLANS DTM format
system(paste0(Fusion, "gridsurfacecreate.exe ", mainDir, outDir, "caldern_GridSurf.dtm ",
              "1 M M 1 32 0 0 ", mainDir, outDir, "classified_GroundPts.las"))

### normalize heights of las point cloud
system(paste0(Fusion, "clipdata.exe", " /height /dtm:", mainDir, outDir,"caldern_GridSurf.dtm ",
              mainDir, inDir, "lidar_files.txt ", mainDir, outDir, "caldern_normalized_point_cloud_LIDAR.las ",
              "476000 5630000 479000 5633000"))

#---------------------------------------------------------------------------#
########### create height slice between 0m (zmin) and 2m (zmax) #############
#---------------------------------------------------------------------------#

### create slice between zmin: 0 and zmax:2#
system(paste0(Fusion, "clipdata.exe"," /zmin:0 /zmax:2 ", mainDir, outDir, "caldern_normalized_point_cloud_LIDAR.las ",
              mainDir, outDir, "caldern_normalized_point_cloud_LIDAR_zmin0_zmax2.las ", "476000 5630000 479000 5633000")) 

###create point count for 1m pixel raster in PLANE dtm
system(paste0(Fusion, "returndensity.exe ",mainDir,outDir, "point_count_all_point_cloud_LIDAR_zmin0_zmax_2_1m.dtm ",
              "1 ", mainDir, outDir, "caldern_normalized_point_cloud_LIDAR_zmin0_zmax2.las"))

###convert to ascii 
system(paste0(Fusion, "dtm2ascii.exe ",mainDir,outDir, "point_count_all_point_cloud_LIDAR_zmin0_zmax_2_1m.dtm ",
              mainDir,outDir, "point_count_all_point_cloud_LIDAR_zmin0_zmax_2_1m.asc"))

#---------------------------------------------------------------------------#
########### calculate CHM ###########
#---------------------------------------------------------------------------#

##canopymodel canopy_surface.dtm
system(paste0(Fusion, "canopymodel.exe ", mainDir, outDir, "canopy_surfaceX.dtm ", "1 M M 1 32 0 0 ",
              mainDir, outDir, "caldern_normalized_point_cloud_LIDAR.las "))
#dtm2tif_canopymodel
system(paste0(Fusion, "DTM2TIF.exe ", mainDir, outDir, "canopy_surfaceX.dtm "))

#---------------------------------------------------------------------------#
########### calculate HOME ###########
#---------------------------------------------------------------------------#

#HOME_median "/raster:p50" , "/gridxy:476000 5630000 479000 5633000 "
system(paste0(Fusion, "gridmetrics.exe ","/raster:p50 ", mainDir, outDir, 
              "caldern_GridSurf.dtm ", "0 ", "1 ", mainDir, outDir, "HOMEgross1.dtm ", mainDir, inDir, "U4765631.las" ))
system(paste0(Fusion, "gridmetrics.exe ","/raster:p50 ", mainDir, outDir, 
              "caldern_GridSurf.dtm ", "0 ", "1 ", mainDir, outDir, "HOMEgross2.dtm ", mainDir, inDir, "U4765632.las"))
system(paste0(Fusion, "gridmetrics.exe ","/raster:p50 ", mainDir, outDir, 
                     "caldern_GridSurf.dtm ", "0 ", "1 ", mainDir, outDir, "HOMEgross3.dtm ", mainDir, inDir, "U4775631.las"))
system(paste0(Fusion, "gridmetrics.exe ","/raster:p50 ", mainDir, outDir, 
                     "caldern_GridSurf.dtm ", "0 ", "1 ", mainDir, outDir, "HOMEgross4.dtm ", mainDir, inDir, "U4775632.las"))


#csv2grid_HOME

system(paste0(Fusion, "CSV2Grid.exe ", mainDir, outDir, "HOMEgross1_all_returns_elevation_stats.csv ", "31 ",
              mainDir, outDir, "HOMExtest1.dtm "))
system(paste0(Fusion, "CSV2Grid.exe ", mainDir, outDir, "HOMEgross2_all_returns_elevation_stats.csv ", "31 ",
              mainDir, outDir, "HOMExtest2.dtm "))
system(paste0(Fusion, "CSV2Grid.exe ", mainDir, outDir, "HOMEgross3_all_returns_elevation_stats.csv ", "31 ",
              mainDir, outDir, "HOMExtest3.dtm "))
system(paste0(Fusion, "CSV2Grid.exe ", mainDir, outDir, "HOMEgross4_all_returns_elevation_stats.csv ", "31 ",
              mainDir, outDir, "HOMExtest4.dtm "))

```


