---
title: "gi-ws-02-1"
author: "Laura Giese, Johannes Schnell, Alena Schmid"
date: "27 Februar 2018"
output: html_document
---

# Deconstruction Work
### description: first try to set up a workflow for deriving vertical forest structure by using LIDAR data 

#### 1. table that describes what attributes of the LIDAR data we want to use for your classification (first try)
```{r, echo = TRUE, eval = TRUE}
Table=data.frame(c("Heterogenität", "Höhe des Kronendachs", "Dichte der Unterschicht", "Kronendeckung/-dichte", "Erhebung", "Terrain Ruggedness", "Hydrologie" ), 
                 c("Punktdichte", "1.Response", "unterhalb 1.Response bis last Response", "1. Response und Punktdichte", "last Response", "last Response", "last response"), 
                 c("1m","1m","1m","1m","1m", "1m", "1m"))
colnames(Table)<-c("Structural Attribute", "Response", "Spatial Scale")                  
print(Table)
```

#### 2. flowchart which illustrates our planned workflow for deriving vertical structural diversity from the LIDAR data
```{r, echo=TRUE, eval = TRUE}
#show flow chart of workflow
#install.packages("png", dependencies=TRUE)
library("png")

setwd("C:/Users/Laura/Documents/Uni/Geoinformationssysteme")
img <- readPNG("plan_s.png")
plot(0:600, type='n', xaxt="n", yaxt="n", xlab="", ylab = "", main = "Flow Chart: Vertical Forest Structure")

lim <- par()
rasterImage(img, 180, -10, 420, 610)
```

#### 3. noticeable points dealing with the difficulties that could occure during the algorithm development:

> Probleme:

1. Boden wird nicht überall erreicht
2. Last response nicht immer = Boden (auch Steine etc.)
3. Lücken im Kronendach auch bei Sträuchern, nicht nur bei Boden
4. Wenn Kronendach sehr dicht, keine Responses unterhalb des Kronendachs
